---
title: "Reproducible document"
author: "Sarah C. Sentis"
format: html
---

## Basics of R

```{r}
1 + 1
```
## Setup code chunk 

```{r setup}
#| message: false
#| warning: false
library(tidyverse)
post_meal_data <- read_csv2(
  # csv2 for ; csv for comma; when underscore e.g. in read_csv, then it is from tidyverse
  here::here("data/post-meal-insulin.csv")
)

small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)

tidier_post_meal_data <- small_post_meal_data |>
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  ) |>
  filter(glykemi == 0) |>
  select(-glykemi)
```

## Showing the data 
```{r}
post_meal_data
```

```{r}
glimpse(post_meal_data)
```

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```
```{r}
#| fig-cap: "Distribution BMI"
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

@fig-bmi-histo


## Exercise: discrete plots 

```{r}
#| fig-cap: "Glykemi Bar-Plot"
#| label: fig-glykemi-bar
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```
@fig-glykemi-bar 




```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = glykemi)) +
  geom_bar()
```

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar()
```

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```

## Putting two plots side by side 

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r}
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

```{r side-by-side figures}
#| fig-cap: "BMI and glycemia, side by side"
#| label: fig-bmi-glykemi
#| fig-subcap:
#| - "Distribution of BMI"
#| - "Number of those with glycemia"
#| layout-ncol: 2

ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()

ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

## Plotting two continuous variables

```{r}
#| fig-cap: "Scatterplot BMI against AUC of blood glucose"
#| label: fig-scatter-bmi-auc-pg
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```

## Pushing changes to remote 

## Pulling changes locally 

## Selecting Data

```{r}
select(post_meal_data, Age)

select(post_meal_data, Age, BMI, Weight)

select(post_meal_data, -Group)

select(post_meal_data, starts_with("PG"))
select(post_meal_data, ends_with("DXA"))
select(post_meal_data, contains("0"))
```

## Renaming specific columns
```{r}
small_post_meal_data |>
  rename(auc_glucose = auc_pg)

small_post_meal_data |>
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  )
```

## Select and rename tasks

```{r}
# Task 1.
post_meal_data |>
  select(contains("OGTT"))

# Task 2.
post_meal_data |>
  select(contains(".."), -contains("...")) |>
  rename(
    P.Glucose.minus5.OGTT = P.Glucose..5.OGTT,
    Insulin.minus5.OGTT.X = Insulin..5.OGTT.X
  )

# Task 3.
basic_info <- select(post_meal_data, id, BMI, Length, Weight, Age)
rename(basic_info, Height = Length)

post_meal_data |>
  select(id, BMI, Length, Weight, Age) |>
  rename(Height = Length)
```


## Filtering data by row 
```{r}
small_post_meal_data |>
  filter(glykemi == 1)

small_post_meal_data |>
  filter(BMI >= 25)

small_post_meal_data |>
  filter(Group == "CTR")
```

```{r}
small_post_meal_data |>
  filter(Group == "CTR", BMI >= 25)

small_post_meal_data |>
  filter(Group == "CTR" | BMI >= 25)
```


## Modify or add columns 
```{r}
tidier_post_meal_data |>
  mutate(
    insulin_glucose_ratio = auc_insulin / auc_glucose,
    Group = if_else(Group == "CTR", "Control", "First-degree relative")
  )
```


## Practice with filter and mutate

```{r}
# Task 1.
post_meal_data |>
  filter(BMI >= 25 & BMI <= 40 & Group == "FDR")

# Task 2.
post_meal_data |>
  mutate(
    percent_body_fat = (Fat.mass.DXA / Weight) * 100
  ) |>
  filter(percent_body_fat > 30)
```



## About me 
- Name: Sarah Sentis 
- Affiliation: University of Southern Denmark 

I am a **Postdoc** and work at the institute of *Biochemistry and Molecular Biology*.
---

## Simple code
```{r}
3 * 3
```

## Testing for Git 
It is raining today and the sun is not shining. 
